---
title: "Maps Investigating"
format: html
runtime: shiny
editor_options: 
  chunk_output_type: inline
---

**Citaiton:**

- Census website containing 2024 total revune for each state based on taxes: https://www.census.gov/data/tables/2024/econ/stc/2024-annual.html

- EPA website containing September 2024 landfill site data: https://www.epa.gov/lmop/landfill-technical-data

```{r echo=FALSE}
#| message: false
#| warning: false

# Initial packages required (we'll be adding more)
library(tidyverse)
library(ggplot2)
library(mdsr)
library(dplyr)
library(tidyr)
library(maps)
library(janitor)
library(data.table)
```

```{r, echo=FALSE}
#| message: false
#| warning: false
TaxOrigional <- read.csv("FY2024_State_taxes.csv")

cleaningtax <- readLines("FY2024_State_taxes.csv")
cleaningtax <- cleaningtax[-c(1:3)]
writeLines(cleaningtax, "FY2024_State_taxes_clean.csv")

FY2024_State_taxes_clean <- read.csv("FY2024_State_taxes_clean.csv")

Taxes <- FY2024_State_taxes_clean

#6tibble(Taxes)

#TaxesA <- Taxes %>%
#  row_to_names(row_number = 1) %>%
#  clean_names()

#tibble(TaxesA)

TaxesB <- transpose(Taxes)

#tibble(TaxesB)

TaxesC <- TaxesB %>%
  row_to_names(row_number = 1) %>%
  clean_names()

#tibble(TaxesC)
```

```{r, echo=FALSE}
#| echo: false
#| message: false
#| warning: false
#| eval: true
TaxesD <- TaxesC |>
  mutate(tax_type = str_to_lower(tax_type))

TaxesD$tax_type <- gsub("\r\n", " ", TaxesD$tax_type)

#tibble(TaxesD)

us_states <- ggplot2::map_data("state")

TaxesE <- TaxesD |>
  mutate(total_taxes = as.numeric(gsub(",", "", total_taxes)))

TaxesE <- TaxesE |>
  mutate(prop_taxes = total_taxes / sum(total_taxes, na.rm = TRUE))

TaxesE |>
  rename(region = tax_type) |>
  right_join(us_states, by = "region") |>
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = total_taxes), color = "black") +
  coord_quickmap()
```

Above is showing a heatmap of total state taxes by state with a lighter blue signifying a higher dollar amount and darker blue signifying a lower amount. Below is a map where you can see the amount of income from property tax in dollars.

```{r, echo=FALSE}
#| message: false
#| warning: false
library(tidyverse)
library(ggplot2)
library(plotly)

us_states <- ggplot2::map_data("state")

TaxesE <- TaxesE |>
  mutate(
    region = tolower(trimws(gsub("\\s+", " ", tax_type)))
  )

df <- TaxesE |>
  right_join(us_states, by = c("region" = "region"))

tax_vars <- c(
  "property_taxes",
  "motor_fuels",
  "tobacco_products",
  "alcoholic_beverages")

plots <- lapply(tax_vars, function(v) {

  ggplot(df, aes(long, lat, group = group)) +
    geom_polygon(aes(fill = .data[[v]], text = region),
                 color = "white") +
    coord_quickmap() +
    theme_void()
})

p <- ggplot(df, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = property_taxes, text = region),
               color = "white") +
  coord_quickmap() +
  theme_void()

gp <- ggplotly(p, tooltip = c("text", "fill"))

value_var <- "property_taxes"

gp <- gp %>%
  layout(
    updatemenus = list(
      list(
        buttons = list(
          list(
            method = "restyle",
            args = list("z", list(df$property_taxes)),
            label = "Property Taxes"),
          list(
            method = "restyle",
            args = list("z", list(df$motor_fuels)),
            label = "Motor Fuels"),
          list(
            method = "restyle",
            args = list("z", list(df$tobacco_products)),
            label = "Tobacco Products"),
          list(
            method = "restyle",
            args = list("z", list(df$alcoholic_beverages)),
            label = "Alcoholic Beverages")))))

gp
```

Next

```{r echo=FALSE}
#| message: false
#| warning: false
Landfills <- read.csv("landfilllmopdata.csv")

Landfills_state <- Landfills |>
  group_by(State) |>
  summarise(n_landfills = n(), .groups = "drop")

state_lookup <- data.frame(
  State = state.abb,
  region = tolower(state.name))

Landfills_state <- Landfills_state |>
  left_join(state_lookup, by = "State")

Landfills_state |>
  right_join(us_states, by = "region") |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n_landfills), color = "black") +
  coord_quickmap() +
  scale_fill_viridis_c(na.value = "white")
```
Above is a density map of landfills across the US as of September of 2024. Below is an interactive map where you can choose to see how many landfills are composing, are owned publicly/privately, and if there is a LFG system in place (LFG = land fill gas, mostly methane created due to anaerobic decomposition of organic material).
```{r, echo=FALSE}
#| echo: false
#| message: false
#| warning: false
#| eval: true

Landfills_clean <- Landfills |>
  mutate(
    composting = ifelse(Composting. == "Yes", 1, 0),
    lfg = ifelse(LFG.Collection.System.In.Place. == "Yes", 1, 0)
  )

state_summary <- Landfills_clean |>
  group_by(State) |>
  summarise(
    composting_total = sum(composting, na.rm = TRUE),
    lfg_total = sum(lfg, na.rm = TRUE),
    ownership_total = n(),  # total sites (or change if you want breakdown)
    .groups = "drop")

state_summary <- state_summary |>
  mutate(
    composting_total = ifelse(composting_total == 0, NA, composting_total),
    lfg_total = ifelse(lfg_total == 0, NA, lfg_total),
    ownership_total = ifelse(ownership_total == 0, NA, ownership_total)
  )

state_lookup <- data.frame(
  State = state.abb,
  region = tolower(state.name))

state_summary <- state_summary |>
  left_join(state_lookup, by = "State")

us_states <- map_data("state")

map_data_full <- us_states |>
  left_join(state_summary, by = "region")

fig <- plot_ly()

# Composting
fig <- fig |> add_trace(
  data = map_data_full,
  type = "scatter",
  mode = "lines",
  x = ~long, y = ~lat,
  split = ~group,
  fill = "toself",
  fillcolor = ~colorRampPalette(c("white","green"))(100)[
    cut(composting_total, breaks = 100, labels = FALSE)],
  line = list(color = "black"),
  name = "Composting",
  visible = TRUE)

# LFG 
fig <- fig |> add_trace(
  data = map_data_full,
  type = "scatter",
  mode = "lines",
  x = ~long, y = ~lat,
  split = ~group,
  fill = "toself",
  fillcolor = ~colorRampPalette(c("white","blue"))(100)[
    cut(lfg_total, breaks = 100, labels = FALSE)],
  line = list(color = "black"),
  name = "LFG",
  visible = FALSE)

# Ownership 
fig <- fig |> add_trace(
  data = map_data_full,
  type = "scatter",
  mode = "lines",
  x = ~long, y = ~lat,
  split = ~group,
  fill = "toself",
  fillcolor = ~colorRampPalette(c("white","red"))(100)[
    cut(ownership_total, breaks = 100, labels = FALSE)],
  line = list(color = "black"),
  name = "Ownership",
  visible = FALSE)

# Dropdown menu
fig <- fig |> layout(
  updatemenus = list(
    list(
      y = 1,
      buttons = list(
        list(method = "update",
             args = list(list(visible = c(TRUE, FALSE, FALSE))),
             label = "Composting"),
        list(method = "update",
             args = list(list(visible = c(FALSE, TRUE, FALSE))),
             label = "LFG Collection"),
        list(method = "update",
             args = list(list(visible = c(FALSE, FALSE, TRUE))),
             label = "Ownership")))))

fig
```
