---
title: "Maps Investigating"
format: html
runtime: shiny
---

**Citaiton:**

- Census website containing 2024 total revune for each state based on taxes: https://www.census.gov/data/tables/2024/econ/stc/2024-annual.html

- Census website containing 2024 voter registration (and if individuals voted) data based on age range and sex: https://www.census.gov/data/tables/time-series/demo/voting-and-registration/p20-587.html
```{r, echo=FALSE}
#| message: false
#| warning: false

# Initial packages required (we'll be adding more)
library(tidyverse)
library(ggplot2)
library(mdsr)
library(dplyr)
library(tidyr)
library(maps)
library(janitor)
library(data.table)
```

```{r, echo=FALSE}
#| message: false
#| warning: false
TaxOrigional <- read.csv("FY2024_State_taxes.csv")

cleaningtax <- readLines("FY2024_State_taxes.csv")
cleaningtax <- cleaningtax[-c(1:3)]
writeLines(cleaningtax, "FY2024_State_taxes_clean.csv")

FY2024_State_taxes_clean <- read.csv("FY2024_State_taxes_clean.csv")

Taxes <- FY2024_State_taxes_clean

#6tibble(Taxes)

#TaxesA <- Taxes %>%
#  row_to_names(row_number = 1) %>%
#  clean_names()

#tibble(TaxesA)

TaxesB <- transpose(Taxes)

#tibble(TaxesB)

TaxesC <- TaxesB %>%
  row_to_names(row_number = 1) %>%
  clean_names()

#tibble(TaxesC)
```





```{r, echo=FALSE}
#| echo: false
#| message: false
#| warning: false
#| eval: true
TaxesD <- TaxesC |>
  mutate(tax_type = str_to_lower(tax_type))

TaxesD$tax_type <- gsub("\r\n", " ", TaxesD$tax_type)

#tibble(TaxesD)

us_states <- ggplot2::map_data("state")

TaxesE <- TaxesD |>
  mutate(total_taxes = as.numeric(gsub(",", "", total_taxes)))

TaxesE <- TaxesE |>
  mutate(prop_taxes = total_taxes / sum(total_taxes, na.rm = TRUE))

TaxesE |>
  rename(region = tax_type) |>
  right_join(us_states, by = "region") |>
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = total_taxes), color = "black") +
  coord_quickmap()
```

Above is showing a heatmap of total state taxes by state with a lighter blue signifying a higher dollar amount and darker blue signifying a lower amount. Below is a map where you can see the amount of income from property tax in dollars.


```{r, echo=FALSE}
#| message: false
#| warning: false
library(tidyverse)
library(ggplot2)
library(plotly)

us_states <- ggplot2::map_data("state")

TaxesE <- TaxesE |>
  mutate(
    region = tolower(trimws(gsub("\\s+", " ", tax_type)))
  )

df <- TaxesE |>
  right_join(us_states, by = c("region" = "region"))

tax_vars <- c(
  "property_taxes",
  "motor_fuels",
  "tobacco_products",
  "alcoholic_beverages")

plots <- lapply(tax_vars, function(v) {

  ggplot(df, aes(long, lat, group = group)) +
    geom_polygon(aes(fill = .data[[v]], text = region),
                 color = "white") +
    coord_quickmap() +
    theme_void()
})

p <- ggplot(df, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = property_taxes, text = region),
               color = "white") +
  coord_quickmap() +
  theme_void()

gp <- ggplotly(p, tooltip = c("text", "fill"))

value_var <- "property_taxes"

gp <- gp %>%
  layout(
    updatemenus = list(
      list(
        buttons = list(
          list(
            method = "restyle",
            args = list("z", list(df$property_taxes)),
            label = "Property Taxes"),
          list(
            method = "restyle",
            args = list("z", list(df$motor_fuels)),
            label = "Motor Fuels"),
          list(
            method = "restyle",
            args = list("z", list(df$tobacco_products)),
            label = "Tobacco Products"),
          list(
            method = "restyle",
            args = list("z", list(df$alcoholic_beverages)),
            label = "Alcoholic Beverages")))))

gp
```


